A New Metropolis-Hastings Test for Fast Minibatch MCMC Inference and Optimization

### 2016-05-16 nips_2016.tex
slight updates to the writeup, added in some comments for 5.1

In this section, we explore how well our algorithm will converge to the posterior distribution.
Define our estimate of $X_{\rm norm}$ as $X_{\rm norm}'$, and let $X_{\xi} = X_{\rm norm} - X_{\rm
norm}'$, so that $\Delta' = \Delta + X_{\rm norm} + X_\xi$.  We also denote $F_X$ as the CDF of $X$.
Let the transition kernel of complete data at sampling step $i$ be $T_i$, and let the transition
kernel of minibatch data at sampling step $i$ is $T_i'$.  We show that the distance of $T_i$ and
$T_i'$ is proportional to our estimate of $X_{\rm norm}$. We also define the stationary distribution
obtained by $T_i$ is $\pi$.  We show the by decreasing the step size, the distribution of our
sampled data converges to $\pi$.

\begin{lemma}
If $|X_\xi| < \zeta$ and $\nabla F_X < k$, the total variance distance between $T_i$ and $T_i'$ is
bounded by $2\zeta k$.
\end{lemma}

@@

Next, we show the transition kernels at each step are bounded if we use the same proposer, i.e.
$\|T_i - T'_i\|_1 \leq 2k\zeta$ with $\| \cdot \|_1$ the standard $L_1$ norm. We can write the
$T_i(\theta_t \mid \theta') = P_a(\theta_t, \theta') q(\theta'\mid \theta_t) +
(1-P_a(\theta_t,\theta')) \delta_D(\theta' - \theta_t)$, where $\delta_D$ is the Dirac delta
function. The distance between two proposal kernels is
\begin{align*}
\|T_i - T'_i \|_1 &= \int_{\theta'} \left| \int_{\theta_t}((P_a-P_a') q(\theta'|\theta_t) + (1-P_a - 1+P_a') \delta_D(\theta_t -\theta')) d\theta_t  \right| d\theta' \\
& = \int_{\theta'} \left| \int_{\theta_t} (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) d\theta_t \right|  d\theta'\\ 
& \leq 2 \zeta k \int_{\theta'} \left| \int_{\theta_t}(q(\theta'|\theta_t) + q(\theta_t|\theta'))  d\theta_t \right| d\theta'= 2 \zeta k,
\end{align*}
as desired.
\end{proof}

@@

\begin{proof}
From Lemma~\ref{lem:gaussian}, only the $\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid
\theta_t))$ term brings randomness into $\Delta'$. To simplify the subsequent notation, we define
$g_i(\theta) = \log p(x_i\mid \theta)$. Then by Taylor's Theorem:
\[
|g(\theta') - g(\theta_t)| = |\nabla g(\theta_t) (\theta' - \theta_t) + o(\theta' - \theta_t)^2| \leq 2\epsilon k.
\]
Thus, with given the $\theta$, and $\theta'$, the variance of $\Delta'$ can be write as:
\begin{align*}
{\rm Var}(\Delta') &= {\rm Var}\left[\sum_{i=1}^n (\log \Pr(x_i | \theta') - \log \Pr (x_i |\theta_t))\right] = {\rm Var}\left[\sum_{i=1}^n g_i(\theta') - g_i(\theta_t)\right] \\
& \leq \sum_{i=1}^n {\rm Var}(g_i(\theta') - g_i(\theta_t)) \leq  4\epsilon^2 m k^2,
\end{align*}
as desired.
\end{proof}

@@

@@

@@

@@

\section{Gaussian Mixture Experiment Details}

@@

\section{Logistic Regression Experiment Details}

@@

\section{Neural Network Experiment Details}

>>>>

In this section, we explore the convergence properties of our algorithm.  Define our estimate of
$X_{\rm norm}$ as $X_{\rm norm}'$, and let $X_{\xi} = X_{\rm norm} - X_{\rm norm}'$, so that
$\Delta' = \Delta + X_{\rm norm} + X_\xi$.  We also denote $F_X$ as the CDF of $X$.  Let the
transition kernel of complete data at sampling step $i$ be $T_i$, and let the transition kernel of
minibatch data at sampling step $i$ is $T_i'$.  We show that the distance of $T_i$ and $T_i'$ is
proportional to our estimate of $X_{\rm norm}$. We also define the stationary distribution obtained
by $T_i$ is $\pi$.  We show the by decreasing the step size, the distribution of our sampled data
converges to $\pi$. We use something called the \emph{total variation distance} between two kernels,
denoted as $\| \cdot \|$.

\begin{lemma}
If $|X_\xi| < \zeta$ and $\nabla F_X < k$, then $\|T_i-T_i'\| \le 2\zeta k$.
\end{lemma}

@@

Next, we show the transition kernels at each step are bounded if we use the same proposer, i.e.
$\|T_i - T'_i\| \leq 2k\zeta$. We write $T_i(\theta_t \mid \theta') = P_a(\theta_t, \theta')
q(\theta'\mid \theta_t) + (1-P_a(\theta_t,\theta')) \delta_D(\theta' - \theta_t)$, where $\delta_D$
is the Dirac delta function. The total variation distance between two proposal kernels is
\begin{align*}
\|T_i - T'_i \| &= \int_{\theta'} \left| \int_{\theta_t}((P_a-P_a') q(\theta'|\theta_t) + (1-P_a - 1+P_a') \delta_D(\theta_t -\theta')) d\theta_t  \right| d\theta' \\
& = \int_{\theta'} \left| \int_{\theta_t} (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) d\theta_t \right|  d\theta'\\ 
& \leq 2 \zeta k \int_{\theta'} \left| \int_{\theta_t}(q(\theta'|\theta_t) + q(\theta_t|\theta'))  d\theta_t \right| d\theta'= 2 \zeta k,
\end{align*}
as desired.
\end{proof}

@@

\begin{proof}
From Lemma~\ref{lem:gaussian}, only the $\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid
\theta_t))$ term brings randomness into $\Delta'$. To simplify the subsequent notation, we define
$g_i(\theta) = \log p(x_i\mid \theta)$. Then by Taylor's Theorem:
\[
|g(\theta') - g(\theta_t)| = |\nabla g(\theta_t) (\theta' - \theta_t) + o(\theta' - \theta_t)^2| \leq 2\epsilon k.
\]
Thus, with given the $\theta$, and $\theta'$, the variance of $\Delta'$ can be write as:
\[
{\rm Var}(\Delta') = {\rm Var}\left[\sum_{i=1}^n g_i(\theta') - g_i(\theta_t)\right] \leq
\sum_{i=1}^n {\rm Var}(g_i(\theta') - g_i(\theta_t)) \leq  4\epsilon^2 m k^2,
\]
as desired.
\end{proof}

@@

{\color{blue}
Daniel: I can borrow some of the description from the Appendix into here, but I need to know what we
are testing. What is the purpose of this experiment? What do we want to show? Once I know that,
writing this part will be easier, and for the Gaussian case in particular, the experiment is pretty
simple and can be run on short notice.

For additional details, see Appendix~\ref{app:gaussian}.
}

@@

{\color{blue}
Daniel: TODO

For additional details, see Appendix~\ref{app:logistic}.
}

@@

{\color{blue}
Daniel: TODO

For additional details, see Appendix~\ref{app:nnet}.
}

@@

\section{Gaussian Mixture Experiment Details}\label{app:gaussian}

@@

\section{Logistic Regression Experiment Details}\label{app:logistic}

@@

\section{Neural Network Experiment Details}\label{app:nnet}
